---
title: "Midterm2"
author: "Saamia"
date: "6/26/2020"
output:
  word_document: default
---

1. 
```{r}
library(rio)
library(tidyverse)
library(data.table)
library(tidyr)
library(dplyr)
library(labelled)
library(doBy)
```

2. Load dataset 
```{r}
inequality_data <- import('inequality.xlsx') 
# checking out the data 
head(inequality_data)
```

3. This dataset is cross-sectional because it is a snapshot from the year 2015. 

4. Sweden = 29.2, Denmark = 28.2 
```{r}
# making subsets of sweden and denmark 
inequality_se <- subset(inequality_data, country == c( 'Sweden'))
inequality_dk <- subset(inequality_data, country == c('Denmark'))
# looking at only inequality_gini
gini_se <- inequality_se$inequality_gini
gini_dk <- inequality_dk$inequality_gini
# printing values
gini_se
gini_dk 
```

5. Brazil = 51.9 
```{r}
# same process as above 
inequality_br <- subset(inequality_data, country == c('Brazil'))
gini_br <- inequality_br$inequality_gini
gini_br
```

6. If Brazil's inequality has decreased significantly then it would seem that a higher score is better. 

7. 
```{r}
# 7. looking at dataframe 
head(inequality_data)
```

8. 
```{r}
# 8. removing accents on Belarus 
accent.remove <- function(s) {
  oldu <- 'ú' #original u with the accent 
  newu <- 'u' #replacement 
  su <- chartr(oldu, newu, s)
  for(i in seq_along(oldu)) su <- gsub(oldu[i], newu[i], s, fixed = TRUE)
  #loop to get rid of the u with accent 
  su #return su 
}
#applying accent.remove 
inequality_data$country =  accent.remove(inequality_data$country)
```

9. Slovenia, Ukraine, Belarus, Czech Republic, Kosovo 
```{r}
# 9. sorting by inequality_gini 
inequality_data = inequality_data[order(inequality_data$inequality_gini),]
# looking at result 
head(inequality_data, 5)
```

10. mean inequality_gini score = 36.81375
```{r echo=TRUE}
# 10. mean gini = 36.81375
# because of missing values need to tell function to ignore those 
mean <- mean(inequality_data$inequality_gini, na.rm = TRUE)
```

11.
```{r}
inequality_data <- 
inequality_data %>%
  mutate(high_inequality = ifelse(inequality_gini > 36.8,
                                  1,
                                  0))
inequality_data <- 
inequality_data %>% 
  mutate(low_inequality = ifelse(inequality_gini < 36.8,
                                  1,
                                  0))
```

12.  
```{r}
# 12. crosstab 
#summaryBy(inequality_gini ~ high_inequality, data = inequality_data, FUN=c(mean, length))
#summaryBy(inequality_gini ~ low_inequality, data = inequality_data, FUN = c(mean, length))
```

13. 
```{r}
# 13. for loop 
actors <- c('World Bank', 'African Development Bank', 'Bill and Melinda Gates Foundation')
for (i in actors){
  print(i)
}
```

14. I chose income share held by lowest 20% because it would help understand the disparity between the richest and poorest/severity of the class divide. 
```{r}
# updating WDI R package 
remotes::install_github('vincentarelbundock/WDI')
library(WDI)
# 14. finding indicator 
WDIsearch('poverty')
```

15. 
```{r}
# 15. importing variable 
povheadcount = WDI(country = 'all', 
                      indicator = c('SI.POV.DDAY'), 
                      start = 2015, end = 2015, extra = FALSE, cache = NULL)
```

16.
```{r}
# 16. renamee variable
setnames(povheadcount, 'SI.POV.DDAY', 'poverty_ratio')
```

17.
```{r}
# 17. merge variable into dataset using left_join to keep rows from inequality_data
merged_df <- dplyr::left_join(x=inequality_data,
                 y=povheadcount,
                 by = c('country', 'year'))
# remove .x .y
merged_df <- 
  merged_df %>%
  select(-c('iso2c.y')) %>%
  rename('iso2c' = 'iso2c.x')
```

18.
```{r}
# 18. remove NAs 
merged_df <- na.omit(merged_df, select = c('inequality_gini', 'poverty_ratio'))
```

19.
```{r}
# 19. filter and piping to keep gini>30 
data_greater_30 <- 
  merged_df %>%
  dplyr::filter(inequality_gini>30)
```

20. 2 
```{r}
grep('ai', data_greater_30)
```

21.  
```{r}
# 21. sum of inequality_gini in data_greater_30 
#sum <- apply(data_greater_30, 2, sum)
#sum
```

22. 
```{r}
# 22. label variables 
var_label(merged_df) <- list(`iso2c` = 'ISO-2 Country Code',
                             `country` = 'Country',
                             `inequality_gini` = 'GINI Score',
                             `year` = 'Year', 
                             `poverty_ratio` = 'Poverty Headcount Ratio')
```

23. 
```{r}
export(merged_df, file = 'final_data.dta')
```

24. 
